Bee visitation, pollination services, and plant yield in commodity and hybrid seed canola

### 2022-09-06 Manuscript/main.tex
Temp switch to Multivac

This study examined how visitation contributes to pollen deposition and seed yield in two canola crops, and showed a strong link between visitation and yield in seed canola, but not in commodity canola.
HB and LCB visitation decreased with distance from their respective hive or shelter, but unexpectedly, LCBs were found more at the edge of the crop.
HB visitation had little direct influence on pollen deposition in either crop type, while LCB visitation had a positive effect on pollen deposition in seed canola.
We found that commodity canola production is largely limited by plant size, while seed canola production is limited by both pollen and plant size.
Finally, our simulation results from the path analyses confirmed this, and showed the strong influence of LCB visitation on seed size and yield.
These results provide more realistic information on the value of pollination services in commodity and seed canola crops, and show how a statistical-mechanistic model of seed production provides richer insite into the process of seed production than linear models alone.

@@

%REVIEW THIS
The limited pollen-transferring behaviour of HBs was also reflected in the pollen counts, as deposition strongly increased with LCB visitation, but not with HB visitation.
LCBs tend to switch between male and female flowers more frequently than HBs do \citep{waytesMsc}, transport more viable pollen between flowers \citep{parker2015}, and tend not to side-work canola flowers \citep{soroka2001}, which may explain their increased pollination efficacy.
Pollen deposition in seed fields also decreased with distance from the edge of the field, and was lower in the centre of the female bays, but this was not explained solely by LCB abundance (Figure \ref{fig:allSEM}).
During foraging bouts, LCBs may travel further between flowers than HBs do \citep{brunet2019}, but pollen carryover from the edge to the centre of the bay female is likely reduced, as LCBs likely visit the edge of the female bay before venturing into the centre \citep{thomson1986, pinnisch1990}.
Pollen counts decreased with distance into the field (independent of visitation), suggests that pollinators are less efficient at transfering pollen.
Foraging \textit{Bombus} take shorter trips between flowers under nectar-rich conditions (area-restricted search, \citealp{pyke1978b, heinrich1979}), suggesting that HBs and LCBs may operate similarly at the centre of seed fields, causing lower pollen deposition at the centre of the field where competition is lower.
% This may be due to lower levels of competition at the centre of the field, leading to a higher nectar (or pollen) standing crop. %LH: more likely that bees at centre of F bays carry less pollen
%LCBs are extremely fast fliers, and are difficult to follow after they leave an inflorescence, meaning that tracking their movements during an entire foraging bout would be almost impossible.
% Visits by HB foragers were almost twice as high in the centre of the female bay, but pollen deposition was about 40\% lower (  Figure \ref{fig:allPollen}), which correspondingly is 
% HBs in seed fields tended to visit more in the centre of the female bay, even close to the edge of the field, yet pollen deposition was still about 40\% lower (Figure \ref{fig:allPollen}).
% While we found that LCBs also tend to avoid the centre of the female bays, and decrease with distance into the field, this extra dependence could have been caused by the fact that the visitation data represents a short ``snapshot'' of the pollen deposition process.
% For example, canola flowers are receptive for 1-2 days before the petals dehisce, while the surveys only recorded 5 or 10 minutes of visitation, meaning we may have not captured the deposition process on the time scale that it occurs, and the extra variance is being accounted for by distance from field edge and the centre of the female bay.
% Leafcutter visitation can be extremely sporadic, so the distance and bay position measurements capture differences in the average visitation process better than 10 minute measurements.
% Pollen deposition also likely changes over the course of the season, as flower availability decreases.

>>>>

We examined how bee visitation contributes to pollen deposition and seed yield in two canola crops, and showed a strong link between visitation and yield in seed canola, but not in commodity canola.
HB and LCB visitation decreased with distance from their respective hive or shelter, but unexpectedly, LCBs were found more at the edge of the crop.
HB visitation had little direct influence on pollen deposition in either crop type, while LCB visitation had a positive effect on pollen deposition in seed canola.
We found that commodity canola production is largely limited by plant size, while seed canola production is limited by both pollen and plant size.
Finally, our simulation results from the path analyses confirmed this, and showed the strong influence of LCB visitation on seed size and yield.
These results provide more realistic information on the value of pollination services in commodity and seed canola crops, and show how a statistical-mechanistic model of seed production provides richer insite into the process of seed production than linear models alone.

@@

%REVIEW THIS
Pollen count models revealed the limited pollen-transferring behaviour of HBs, as deposition strongly increased with LCB visitation but not HB visitation.
LCBs tend to switch between male and female flowers more frequently than HBs do \citep{waytesMsc}, transport more viable pollen between flowers \citep{parker2015}, and tend not to side-work canola flowers \citep{soroka2001}, which may explain their increased pollination efficacy.
Pollen deposition in seed fields also decreased with distance from the edge of the field, and was lower in the centre of the female bays, but this was not explained solely by LCB abundance (Figure \ref{fig:allSEM}).
During foraging bouts, LCBs may travel further between flowers than HBs do \citep{brunet2019}, but pollen carryover from the edge to the centre of the bay female is likely reduced, as LCBs likely visit the edge of the female bay before venturing into the centre \citep{thomso=n1986, pinnisch1990}.
Pollen counts decreased with distance into the field (independent of visitation), suggests that pollinators are less efficient at transfering pollen.
Foraging \textit{Bombus} take shorter trips between flowers under nectar-rich conditions (area-restricted search, \citealp{pyke1978b, heinrich1979}), suggesting that HBs and LCBs may operate similarly at the centre of seed fields, causing lower pollen deposition at the centre of the field where competition is lower.
% This may be due to lower levels of competition at the centre of the field, leading to a higher nectar (or pollen) standing crop. %LH: more likely that bees at centre of F bays carry less pollen
%LCBs are extremely fast fliers, and are difficult to follow after they leave an inflorescence, meaning that tracking their movements during an entire foraging bout would be almost impossible.
% Visits by HB foragers were almost twice as high in the centre of the female bay, but pollen deposition was about 40\% lower (  Figure \ref{fig:allPollen}), which correspondingly is 
% HBs in seed fields tended to visit more in the centre of the female bay, even close to the edge of the field, yet pollen deposition was still about 40\% lower (Figure \ref{fig:allPollen}).
% While we found that LCBs also tend to avoid the centre of the female bays, and decrease with distance into the field, this extra dependence could have been caused by the fact that the visitation data represents a short ``snapshot'' of the pollen deposition process.
% For example, canola flowers are receptive for 1-2 days before the petals dehisce, while the surveys only recorded 5 or 10 minutes of visitation, meaning we may have not captured the deposition process on the time scale that it occurs, and the extra variance is being accounted for by distance from field edge and the centre of the female bay.
% Leafcutter visitation can be extremely sporadic, so the distance and bay position measurements capture differences in the average visitation process better than 10 minute measurements.
% Pollen deposition also likely changes over the course of the season, as flower availability decreases.
